\documentclass[pdftex,letterpaper,11pt]{article}%
\usepackage[export]{adjustbox}
\usepackage{multirow}
\usepackage[arabicsections]{dpugatex}
%\usepackage{dpppl}
%\usepackage{lscape}
\usepackage{pdflscape}

\usepackage{tikz}

\usepackage[british]{babel}
\usepackage{amssymb,amsmath,warpcol,url,ctable,multirow,caption,threeparttable,float,soul,gensymb}
%\usepackage[noend]{algpseudocode}
\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother


\usepackage{subcaption}


\usepackage{xcolor,colortbl}
\definecolor{verylightgray}{gray}{0.90}
\newcommand{\verylightgray}[1]{\cellcolor{verylightgray}#1}


\usepackage{arydshln}
\setlength\dashlinedash{1.5pt}
\setlength\dashlinegap{2.5pt}
\setlength\arrayrulewidth{0.8pt}

\usepackage[utf8]{inputenc}
\usepackage{eurosym}
\usepackage{soul}
\setstcolor{red}
\definecolor{lightgreen}{rgb}{0.564706,0.933333,0.564706}
\soulregister\citet7 % These \soulregister commands allow us to avoid problems between soul functions (\hl, \st, ...) and these other functions (\citet,\citep,\ref,\footnote, ...). Include more if it is necessary.
\soulregister\citep7
\soulregister\cite7
\soulregister\ref7
\soulregister\footnote7
\soulregister\textit7
\soulregister\textbf7
\soulregister\textsc7
\soulregister\`7
\soulregister\'7
\soulregister\% 7
\usepackage{tabularx,tabulary}
\usepackage[osf]{mathpazo}
\usepackage[abs]{overpic}

\usepackage{hyperref}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\hypersetup{colorlinks=true, breaklinks=true, citecolor= darkblue, linkcolor=blue, urlcolor=blue}

\usepackage{chngcntr}

\newcommand{\mc}[3]{\multicolumn{#1}{#2}{#3}}
\newcommand{\mcc}[1]{\multicolumn{1}{c}{#1}}
\newcommand{\mcb}[1]{\multicolumn{1}{c}{\bf #1}}

\newcommand{\mr}[3]{\multirow{#1}{#2}{#3}}
\newcommand{\mg}[3]{\multicolumn{#1}{#2}{\verylightgray #3}}


\onehalfspacing%
%\doublespacing%

\newenvironment{tablenote}[1]{\begin{list}{}{\vskip-5mm\relax
\setlength{\leftmargin}{#1} \setlength{\rightmargin}{\leftmargin}}
\item[]\footnotesize\vskip-7pt
{\em Notes}:\space\ignorespaces}{\end{list}}

\newcommand{\jdlradded}[1]{#1}
\newcommand{\dpadded}[1]{#1}
\newcommand{\jdlrdeleted}[1]{}
\newcommand{\dpdeleted}[1]{}
\newcommand{\jdlrcomment}[1]{}
\newcommand{\dpcomment}[1]{}
\newcommand{\comment}[1]{}
\newcommand{\martin}[1]{{\color{blue} Martin: [{#1}]}}
\newcommand{\dani}[1]{{\color{purple} Dani: [{#1}]}}

\begin{document}
\begin{titlepage}
\vspace*{1ex}
\begin{minipage}{\textwidth}
\begin{center}%

    {\textsb{\LARGE Classifying urban form at national scale\\ \textit{The British morphosignatures}}}\\[4ex]%


{\Large\textbf{Martin Fleischmann}\footnote[1]{THANKS.}\footnote[2]{Geographic
Data Science Lab, Department of Geography and Planning, University of
Liverpool, Roxby Building , 74 Bedford St S , Liverpool , L69 7ZT, United
Kingdom}\footnote[4]{E-mail: \url{D.Arribas-Bel@liverpool.ac.uk}; phone: +44 (0)151 795 9727; website: \url{http://darribas.org}.
}}\\[1mm]
%
{\Large\textsb{Daniel Arribas-Bel}\footnotemark[1]\footnotemark[2]\footnote[3]{
E-mail: \url{M.Fleischmann@liverpool.ac.uk}; website: \url{https://martinfleischmann.net/}.}}\\[1mm]
{\large\textit{Geographic Data Science Lab, University of Liverpool} }\\[2.5ex]
%
July 2021\vspace{1.5ex}
%
\end{center}
%
\begin{abstract}
In this paper, we use the recent spatial signatures framework to develop
\textit{morphosignatures}, a characterisation of space based on urban form,
and we apply them to the whole of Great Britain.
%
Identification of recurring patterns in the built environment is deeply embedded within
all schools of urban morphology.
%%
A long-standing challenge in this context has been the ability to scale
the analysis to cover large regions. In effect, advancing the
systematic study of urban morphology has been limited to the ability to deploy
granular methods to large portions of the urban landscape.
%%
The recent rise of
morphometrics, and its ability to scale while retaining detail, has given
urban morphology a fundamentally new toolkit to analyse the form of urban
fabric at metropolitan or even national extents.
%
In this tradition, morphosignatures are conceptually defined as an aggregation of granular elements into
contiguous areas based on similar nature. We adopt
the enclosed tessellation cell (ETC), the result of combining street networks
and other urban “delimiters” with building footprints, as the core spatial unit.
For each ETC, we calculate a wide range of characters that capture aspects
of their spatial organisation, from dimension and shape of individual
features to their spatial distribution, intensity or connectivity.
We then group these ETCs using the
K-Means algorithm, effectively deriving a comprehensive typology of urban form.
%
We employ this approach on the case of Great Britain, illustrating both the potential and
limits of the analysis of urban form at scale. Our resulting classification identifies
19 types of morphosignatures that can be organised into three macro
groups: countryside, suburban low density development, and dense city centres.
%
The main
contribution of the proposed method, with respect to traditional morphological studies,
resides in the scalability of the analysis while retaining granularity. We
argue national-scale classification such as that of the British
morphosignatures allows us to ask fundamentally new questions and provide
complementary answers to those asked in smaller scale studies.

\end{abstract}
%
\vspace{1.5ex}
%
\vspace*{-1.5ex}
%
\end{minipage}
\end{titlepage}

\input{01_intro.tex}

\input{02_method.tex}

\input{03_results.tex}

\input{04_discussion.tex}

\clearpage




\bibliographystyle{apalike}
\bibliography{refs}

\clearpage

\end{document}
